---
title: "Week1 - Getting and Cleaning Data"
author: "Sumit"
date: "8/17/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## what are we going to cover in this course: 
One of the major components of a data scientist's job is to collect and clean data. 
In this course we will learn to collect data from various sources and various formats.
This includes sources such as:--
* Data Collection
        + Raw files (csv, xlsx)
        + Databases (MySQL)
        + APIs

* Data Formats
        + Flat files (csv, txt)
        + XML
        + JSON

* Making data tidy
* Distributing data
* Scripting for data cleaning

## Raw Data vs Processed Data
Raw data is taken from the source and often hard to use for analysis. Raw data 
needs to be processed so that it can be used for analysis. Processing may involve
merging, subsetting, transforming etc. Each process should be recorded. 

## The components of the tidy data
#### What we need to have after we convert the data from Raw data to Tidy data
1. Raw Data
2. Tidy data 
3. Meta data i.e. code book describing each variable and its values in Tidy data
4. An explicit and exact recipe we used to go from step 1 to 2

#### What a raw data may look like
1. A strange binary file your computer spits out
2. The unformatted excel with 10 worksheets
3. A complicated JSON data we got from scrapping the Twitter API

#### Raw data Don'ts:--
1. Run any software on the data
2. Manipulate anything 
3. Remove anything from the dataset
4. Summarize the data 

#### How should a Tidy data look like
1. Each variable that you measure should be in one column
2. Each observation should be in different row
3. There should be one table for each kind of variable
4. If you have multiple tables, there should be a unique ID connecting them. 

#### The code book shall have information about :--
1. variables (including units)
2. Summary choices (mean / median revenue)
3. Experimental study design 

#### Converting the raw file into a tidy dataset:--
1. This should ideally be a script
2. The input to this script shall be the raw data and the output shall be the tidy data
3. There should be no parameters in this script and everything shall be fixed

In some cases where it's not possible to script every script, you should provide 
clear instructions like - 
* Step 1: Take the raw file and summarize ...
* Step 2: ...

## Downloading data
### Checking and creating directories
**file.exist(directory name)** this will check if the directory exist  
**dir.create(directory name)** this will create a directory

```{r}
file.exists("tempdir")
```

### Getting the data from internet
**download.file(*<url>*)** this will download file from internet

For this section, we will be using data from Baltimore traffic data. 
The link for this is given below:

[https://data.baltimorecity.gov/datasets/fixed-speed-cameras/explore?location=39.300050%2C-76.582539%2C11.91](https://data.baltimorecity.gov/datasets/fixed-speed-cameras/explore?location=39.300050%2C-76.582539%2C11.91)

### Other places to get the datasets for analysis
1. [https://www.kaggle.com/datasets](https://www.kaggle.com/datasets)  
2. [https://www.kaggle.com/kkhandekar/covid-vaccination-stats-in-new-zealand?select=Cumulative.csv](https://www.kaggle.com/kkhandekar/covid-vaccination-stats-in-new-zealand?select=Cumulative.csv)  
3. [Free public datasets for analysis](https://www.tableau.com/learn/articles/free-public-data-sets)  
4. [AWS Open Datasets](https://registry.opendata.aws/)  
5. [XLSX data](https://www.datarefuge.org/dataset?res_format=XLSX)  

## Downloading process

```{r}
fileUrl <- "https://www.kaggle.com/kkhandekar/covid-vaccination-stats-in-new-zealand?select=Cumulative.csv"

if (!file.exists("data")){
        dir.create(".\\data")
        print('New directory data created')
} else print('Directory data already exist')
```

```{r}
setwd(".\\data")
download.file(fileUrl, destfile = 'covidcase.csv', method = 'curl')
dateDownload = date()
list.files('.\\')
dateDownload
```

# Reading local files
You can read local files using read.table() and read.csv() functions

## Reading CSV files

```{r}

covidData <- read.table('.\\data\\Cumulative.csv', sep = ',', header = TRUE)
head(covidData, 5)

```

You can also use read.csv files if you have a csv files

```{r}
covidData <- read.csv(".\\data\\Cumulative.csv")
head(covidData, 5)
```

##### Parameters you can pass along with read.csv()
* quote-- set quote as "'" which means there are quotes single quote in the data
* na.strings-- any specific things defining na's like 9, -99 etc. 
* nrows-- how many rows we want to read
* skip-- how many line we have to skip before statrting to read

## Read XLSX files

**Downloading the file using the same download.file() function, which is agnostic
of the file type.** 
```{r}
fileUrl <- "https://www.datarefuge.org/dataset/e26d379e-05d9-4ed3-8d1f-9e44738149d8/resource/bd3902e5-cf0f-4198-96de-8c48eb06563e/download/internet_enabled_mobile_phone_use_by_gender.xlsx"

download.file(fileUrl, destfile = '.\\data\\mobileUsage.xlsx', method = 'curl')
dateDownload <- date() 
list.files(".\\data\\")
dateDownload
```

Read the files using read.xlsx() / read.xlsx2()  function.   
These functions can be found in **XLSX Package**

<span style="color: red;">**If there are issues in running the xlsx function, please refer to the** 
**the link below** </span>

[XLSX does't work](https://www.coursera.org/learn/data-cleaning/discussions/forums/pYncWij2Eea8jw6UvTi2Tw/threads/b5TU4RX-EeawWRJ2-Hqumw/replies/EEX_OBYMEea44Qp49aJR6w)

1. Install 3 packages: xlsx, rJava, xlsjars
2. Install the right version (32/64bits) of Java at http://java.com/en/download/win10.jsp
3. Load the 3 packages using library(). 


```{r}
library(xlsx)
dat <- read.xlsx(".\\data\\cumulative_xl.xlsx", sheetIndex = 1, header = T )
head(dat)
```

if you only want to select a few columns and few rows, we can use parameters like 
colindex and rowIndex

```{r}
library (xlsx)
subData <- read.xlsx(".\\data\\cumulative_xl.xlsx", sheetIndex = 1, header = T,
                     colIndex = 1:3, rowIndex = 1:5)
subData
```
## Other useful notes
* We can use *write.xlsx* function to write in excel
* *read.xlsx2* is faster in reading excel files but can be unstable when reading
subset of rows
* **XLConnect** package is very useful if we are doing heavy excel functions.
* **XLConnect vignette** is a great place to start to learn about these packages
* We should save our data in a database or in a csv, tab / txt file 








