---
title: "Data.Table package"
author: "Sumit Kamra"
date: "8/23/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data.table package  
The data.table package is extremely fast way to process the data as the code is
written in C. This comes with slightly difficult to remember syntax though. 

* All functions that accepts data frame also work with data.table
* Much faster at subsetting, group and updating.

## Creating tables just like data frame
```{r}
set.seed(10)
df <- data.frame(x = rnorm(9), y = sample(1:100, 9), 
                 z = rep(c('a','b','c'), each = 3))
head(df, 3)
```

```{r}
library(data.table)
set.seed(10)
dt <- data.table(x = rnorm(9), y = sample(1:100, 9), 
                 z = rep(c('a','b','c'), each = 3))
head(dt, 3)
```
## Checking how many tables have been created
```{r}
tables()
```
It tell us which all data tables have been created, how many rows, columns, and 
if there is any key. 

## subsetting the data tables (same as data frames)

```{r}
dt[2,] 
```
```{r}
dt[dt$z=='a',c('x', 'z')]
```

## Subsetting rows 
If we only specifiy one element in while subsetting, it subsets the rows. 
```{r}
dt[c(2,3)]
```
## This is different for data frame where columns subset 
```{r}
df[c(2,3)]
```
## Let's try to subset columns in the data table
```{r}
dt[,c(2,3)]
```
## Columns subsetting in data.table
* The subsetting function is modified for data table
* The argument you pass after a comma is called an expression
* In R, the expression is a collection of statement enclosed in a curly brackets

```{r}
{
        x = 1
        y = 1
}
k = {print(10); 5}

```

```{r}
print(k)
```
## Calculating values for variables with expressions
in data tables, for columns, you can pass a list of functions that you want to 
apply on the variables. In the example below, we will calculate a mean of x and 
sum of y in the data table object - dt
```{r}
dt[,list(mean(x), sum(y))]
```

.. And here we are producing a table of variable z 
```{r}
dt[,table(z)]
```
## Adding new columns: 

```{r}
dt[, pr:=x*y]

dt

```
## Multiple step operation 

```{r}
# Removing the variable pr
dt[, pr:= NULL ]
# Adding a new variables m through multiple operation.
dt[, m:= {tmp <- x+y; log(tmp+5)}]
dt
```
## Plyr like operations
Adding new variables based on conditions. In this example, we will add a column
to indicate if y is great or less than 50. 

```{r}
dt[,m:= NULL]
dt[, y_lessthan_50:= y<50 ]
dt
```
## Adding a column with a condition
```{r}
dt[,y_lessthan_50:= NULL]
dt[, y_status:= NULL]
dt[,y_status:= { 
        tmp = y<30 ; tmp
        }
   ]
dt
```
The experiment above didn't work

## Plyr like operations - group by (by keyword)

```{r}
dt[, m:= mean(y), by=y_status]
dt

```
it calculates aggregated mean of y, aggregated over the variable we use in 'by'

## Special variables
.N is a special variables which is an integer
* It contains no of times that a group appear in the data table

Now I am creating a data table with multiple nos of letter
```{r}
lt <- data.table(x = sample(c("a", "b", "c"), 1E5, replace = T))
```

Now I am using .N grouped by x
```{r}
lt[,.N, by=x]

```
## Keys
Unique aspect of data tables is that you can set keys which speeds up the process
of subsetting, sorting etc. 

```{r}
ndt <- data.table(x = sample(c ("a", "b", "c"), 100, T), y=rnorm(100))
setkey(ndt, x)
head(ndt["a"], 10)
```
Showing only the rows with x = "a" because x is assigned as the key. 

## Joins
Setting key is immensly useful when joining the two data tables. 
The example below

```{r}
dt1 <- data.table(x=c("a","b", "c"), y=11:13)
dt2 <- data.table(x=c("a", "m", "c", "b"), z=1:4)
setkey(dt1, x)
setkey(dt2, x)
merge(dt1, dt2)

```



