---
title: "Reading XML JSON Files"
author: "Sumit Kamra"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading XML Files

### Reading the XML file and finding it's nodes
The code below uses XML library, reads an XML file and find the root nodes.   
In this example below, I had to use the GET function as directly passing the
url in xmlUrl didn't work. 

```{r}
library(XML)
library(httr)
fileUrl <- "https://www.w3schools.com/xml/simple.xml"
r = GET(fileUrl)
doc <- xmlTreeParse(r, useInternalNodes = TRUE, isURL = TRUE)
rootNode <- xmlRoot(doc)
xmlName(rootNode)

```
### Accessing the part of XML file
#### The first food node
```{r}
rootNode[1]

```

```{r}
rootNode[[1]]
```

#### Accessing subnodes 
In this example, we are accessing the second subnode with first node. We can also 
access the content through xmlValue function:--- 
```{r}
rootNode[[1]][[2]]

xmlValue(rootNode[[1]][[2]])
```
This can also be performed on all the nodes with an xmlSApply function 

```{r}
xmlSApply(rootNode[[1]], xmlValue)
class(xmlSApply(rootNode[[1]], xmlValue))
```

While using xmlApply function would return a list
```{r}
xmlApply(rootNode[[1]], xmlValue)
class(xmlApply(rootNode[[1]], xmlValue))
```
# Using XPath to extract elements
#### XPath is an entirely different language
* /node - top level node
* //node - node at any level 
* node[@attr-name] - node with an attribute name
* node[@attr-name ='bob] - node with attribute name = 'bob' 

The example below extracts name, price and description of rootNode and turns it
into a data frame. 

```{r}
name = xpathSApply(rootNode, '//name', xmlValue)
price = xpathSApply(rootNode, '//price', xmlValue)
desc = xpathSApply(rootNode, '//description', xmlValue)
d<- data.frame(
        name, price, desc
)
d
```
```{r}
fileUrl <- "https://www.espn.com/nfl/team/stats/_/name/bal"
r<- GET(fileUrl)
doc <- htmlTreeParse(r, useInternalNodes = TRUE)
rootNode <- xmlRoot(doc)

names(rootNode)


```



