---
title: "Reading API"
author: "Sumit Kamra"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Useful references: 
* [introduction to web scraping](https://nceas.github.io/oss-lessons/data-liberation/intro-webscraping.html)
* 

## API
API is Application Programming Interface.   

[**Github tutorial for accessing Github API**](https://github.com/r-lib/httr/blob/master/demo/oauth2-github.r)

[**Github httr tutorial**](https://github.com/r-lib/httr)

Need to covered later through tutorials. Refer to the file named 
JHU lecture - readingFromAPIs.pdf

## Data from other sources

If you are looking to pull data from other sources (and there are tons of souces)
In general the best way to search R package is to use query like this: 
*data store mechanism R Package*  

For example:  
*MySQL R Package*  

## Functions to manipulate connections

```{r connections}
?connections
```

**ALWAYS REMEMBER TO CLOSE A CONNECTION**

## Foreign package
Functions used to access files from other software

General Syntax:--- read.xxx (xxx is file extension of other software)
* read.arff(Weka Files)
* read.dta (Stata)
* read.mtp (Minitab)
* read.octave (Octave)
* read.spss (SPSS)
* read.xport (SAS)

**For more information please click use the link below**  
["https://cran.r-project.org/web/packages/foreign/foreign.pdf"]("https://cran.r-project.org/web/packages/foreign/foreign.pdf")

## Reading from other sources such as:  
* Databases
* Images (jpeg, png, EBIimage)
* GIS Data
* tuneR / seewave for musical data

**Please refer to the file JHU lecture - readingFromOtherSources.pdf** 











