---
title: 'Week 2: Getting Data from Various Sources'
author: "Sumit Kamra"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing RMySQL package

[Installation instructions used](https://www.ah-ruhe.de/2013/07/23/installing-rmysql-under-windows/)

## Connecting with MySQL database

[UCSC Genome MySQL Page](https://genome.ucsc.edu/goldenPath/help/mysql.html)

```{r}
library(RMySQL)
uscDb <- dbConnect(MySQL(), user = "genome",
                host = "genome-mysql.soe.ucsc.edu")
result <- dbGetQuery(uscDb, "show databases;")
dbDisconnect(uscDb)

```
```{r}
head(result)
```
## Access a particular database "hg19" from the genome databases

```{r}
hg19 <- dbConnect(MySQL(), user = "genome", db = "hg19",
                  host = "genome-mysql.soe.ucsc.edu")
allTabs <- dbListTables(hg19)
length(allTabs)

```
#### Listing all fields in hg19 database
```{r}
allTabs[1:10]
```
#### List all fields in a particular table in hg19 database
```{r}
dbListFields(hg19,"affyExonProbeAmbiguous" )
```
#### Querying a table from the database
```{r}
dbGetQuery(hg19, "select count(*) from affyExonProbeAmbiguous")

```

#### Read from the table using dbReadTable command
```{r}
data <- dbReadTable(hg19, "affyExonProbeAmbiguous")
head(data)

```
## Selecting a subset from a table 
#### Use dbSendQuery command to extract a subset of a table

```{r}
query <- dbSendQuery(hg19, 
        "select * from affyExonProbeAmbiguous where bin between 500 and 1500")
dat <- fetch(query)
head(dat)

```
#### Use fetch 
```{r}
datSmall <- fetch(query,n=10)
dim(datSmall)
dbClearResult(query)
```
```{r}
dbDisconnect(hg19)
```

## Important references

[**Click here to know more about advanced DBI usage**](https://cran.r-project.org/web/packages/DBI/vignettes/DBI-advanced.html)

[MySQL commands](https://www.pantz.org/software/mysql/mysqlcommands.html)

[R Bloggers MySQL and R](https://www.r-bloggers.com/2011/08/mysql-and-r/)


