---
title: "Reading Web Data"
author: "Sumit Kamra"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Webscraping
Reading data programmatically from the web is called web scraping. 

The link here is extremely interesting. [How Netflix reverse engineered hollywood ](https://www.theatlantic.com/technology/archive/2014/01/how-netflix-reverse-engineered-hollywood/282679/)

#### We will be using the Google Scholar link here: 

[Google Scholar - Jeff Leak Page](https://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en)

In the code below, we wil open a connection to a url and read out a few lines
from the web page. 
```{r setconnection}
con <- url('https://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en')
htmlText <- readLines(con, 2)
close(con)


```

#### Using the get command fro httr package
```{r useGET}
library(httr)
library(XML)
urlText <- 'https://scholar.google.com/citations?user=HI-I6C0AAAAJ&hl=en'
con <- GET(urlText)
htmlText <- content(con, as="text")
parsedHtml <- htmlParse(htmlText, asText = T)
xpathSApply(parsedHtml, "//title", xmlValue)

```
```{r GETauthicatedapps}
library(httr)
pg2 <- GET('https://httpbin.org/basic-auth/user/passwd',
           authenticate("user", "passwd"))
pg2

```
```{r listing_names}
names(pg2)
```
#### Usig handles: 
If you assign a handle once, you can use it to get the webpage through GET
multiple times, and if you authticate once, the cookies will stay and you 
wont need to authenticate again. 

```{r using_handles}
google <- handle("http://google.com")
pg1 <- GET(handle=google, path = "/")
pg2 <- GET(handle=google, path = "search")
pg1
pg2
```
## Tips: 
* Use r-bloggers and search for webscraping [Link](https://cse.google.com/cse?cx=005359090438081006639%3Apaz69t-s8ua&ie=UTF-8&q=web+scraping&sa=Go)
* 

