---
title: "Editing Text Variables"
author: "Sumit Kamra"
date: "9/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Summary of the functions covered in the module
* tolower()
* toupper()
* strsplit()
* sub() 
* gsub()
* grep()
* grepl()
* nchar()
* substr()
* paste()
* paste0()


## Editing text variables

```{r}
cameraData <- read.csv("Fixed_Speed_Cameras.csv")
names(cameraData)
```
#### Modifying the names of the variables
```{r}
tolower(names(cameraData))
```
```{r}
toupper(names(cameraData))
```
#### Removing dots from variable names
```{r}
names(cameraData)
```
```{r}
splitNames <- strsplit(names(cameraData), "\\.")
splitNames[[1]]
```
```{r}
firstElement <- function(x) x[[1]]
names(cameraData) <- sapply(splitNames, firstElement)
names(cameraData)[[1]] <- "i"
names(cameraData)
```

```{r}
library(dplyr)
cameraData <- rename(cameraData, id_name = i, fid_id = fid)
names(cameraData)
```
#### Removing the special characters from the variable names
```{r}
sub("_", "", names(cameraData))
```
#### sub only replaces first underscore
```{r}
testName <- "this_is_my_name"
sub("_", "", testName)   ## Replaces first underscore
gsub("_", "", testName)  ## Replaces all underscores

```
## Searching for specific values in variables

```{r}
## Looking for Ave
grep("Ave", cameraData$street)

table(grep("Ave", cameraData$street))

table(grepl("Ave", cameraData$street))

```
#### Can also subset by grepl
```{r}
cameraData[grepl("Ave", cameraData$street),]
```

```{r}
grep("Ave", cameraData$street, value = TRUE)
```
```{r}
grep("JeffStreet", cameraData$street)
length(grep("JeffStreet", cameraData))
```
## Library stringr
```{r}
library(stringr)
nchar("Sumit Kamra")

```

```{r}
substr("Sumit Kamra", 2, 4) ## both indices are inclusive
```

```{r}
paste("Sumit", "Kamra", sep = "-")
paste0("Hi", "How", "are")
```


