---
title: "Creating New Variables"
author: "Sumit Kamra"
date: "8/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why create new variables
* Missingness indicators
* "cutting up" the quantitative data to create factors from numeric data
* Applying transformation

#### We will be using the baltimore restaurants data for this exercise 
```{r}
resData <- read.csv("Restaurants.csv")
```

#### Using sequencing to create indices for your datasets

```{r}
s <- seq(1,10, by=2); s
```
```{r}
s<- seq(1,10, length = 4); s
```
```{r}
s<-seq( along = resData$Y); 
str(s)
```

#### Subsetting variables

```{r}

resData$nearme <- resData$nghbrhd %in% c("Roland Park", "Homeland")
table(resData$nearme)
```
#### Creating binary variables

```{r}
resData$lessThanHundred <- ifelse(resData$fid <100, T, F)
table(resData$lessThanHundred, resData$fid <100)
```
#### Creating categorical variables
Categorical variables are created from numeric variables through cut function

```{r}
resData$fidgroups <- cut(resData$fid, breaks = c(0, 20,40,60,80,100))
table(resData$fidgroups, useNA = "ifany")
```


```{r}
resData$fidgroups <- cut(resData$fid, breaks =  quantile(resData$fid))
table(resData$fidgroups, useNA = "ifany")
```
#### Another way of doing it through Hmisc library
```{r}
# Use Hmisc library and cut2 function 
library(Hmisc)
resData$fidgroups <- cut2(resData$fid, g = 4)
table(resData$fidgroups, useNA = "ifany")
```
```{r}
resData$zip <- as.numeric(resData$zipcode)
resData[1157, "zip"] <- 21226

```
Using the cut function again on numeric zip
```{r}
resData$zipgroups <- cut2(resData$zip, g=5)
table(resData$zipgroups, useNA = "ifany")
```
```{r}
table(resData$zip, resData$zipgroups, useNA = "ifany")
```
#### creating facor variables
zip is a numeric variables and we can convert them into factor variable by using 
factor function

```{r}
resData$zcf <- factor(resData$zip)
resData$zcf[1:10]
```
#### Some more useful stuff ralated to factors
```{r}
set.seed(20)
yesno <- sample(c('yes','no'), size = 20, replace = T)
yesnofac <- factor(yesno, levels = c('yes' , 'no'))
yesnofac
```
You can also convert factors into numeric values if you like to use in some kind
of model - it's very useful
```{r}
as.numeric(yesnofac)

yesnofac
```
#### Using mutate function to mutate and add factor variable to the new data set

```{r}
library(Hmisc); library(plyr)
resData2 <- mutate(resData, zipGroups = cut2(zip, g=4))
table(resData2$zipGroups, useNA = "ifany")
                   
```

## Other useful function for transformation

abs(x)	absolute value
sqrt(x)	square root
ceiling(x)	ceiling(3.475) is 4
floor(x)	floor(3.475) is 3
trunc(x)	trunc(5.99) is 5
round(x, digits=n)	round(3.475, digits=2) is 3.48
signif(x, digits=n)	signif(3.475, digits=2) is 3.5
cos(x), sin(x), tan(x)	also acos(x), cosh(x), acosh(x), etc.
log(x)	natural logarithm
log10(x)	common logarithm
exp(x)	e^x

## Further reference

[https://www.statmethods.net/management/functions.html](https://www.statmethods.net/management/functions.html)
[plyr tutorial](https://plyr.had.co.nz/09-user/)



