---
title: "dplyr function"
author: "Sumit Kamra"
date: "8/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## dplyer function

It is used to manipulate data frames. It can be used with data.tablble package 
and rdbms. 

#### Grammer of data manipulation
* select: returns a subset of columns of the data frame
* filter: returns a subset of the data frame
* arrange: reorder rows 
* rename: rename variables
* mutate: adds columns or transforms existing 
* summarise: generate summary statistics

## Basic toold - dplyr package

```{r}
library(dplyr)
# Downloading chicago.rds file from github

chicago <- readRDS("chicago.rds")
```

### Select function
```{r}
head(select(chicago, city:date))
```

```{r}
head(select(chicago, -(city:dptp)))
```

## filter function
```{r}
chic.f <- filter(chicago, (pm10tmean2 > 30 & dptp <30))
head(chic.f)
```
## arrange function
```{r}
chic.s <- arrange(chicago, desc(date))
tail(chic.s)
```


## renam function
```{r}
chicago <- rename (chicago, pm25 = pm25tmean2, dewpoint=dptp)
names(chicago)
```
## mutate function
```{r}
chicago <- mutate(chicago, pm25trend = pm25 - mean(pm25, na.rm = TRUE))
tail(select(chicago, pm25, pm25trend))
```

#### Creating factor variables
```{r}
chicago <- mutate(chicago, tempcat = factor(1 * (tmpd >80), labels = 
                                                    c("Cold", "Hot")))
head(select(filter(chicago, tmpd>80), tmpd, tempcat))
```

#### Using group_by
```{r}
hotcold <- group_by(chicago, tempcat)
hotcold
```

#### Summarizing data after group_by

```{r}
summarise(hotcold, pm25_mean=mean(pm25, na.rm = T), 
          pm10_median= median(pm10tmean2, na.rm = T))
```
#### Summarising by year
1. Mutate to add a categorical year variable in the data
2. Group by the year variable
3. summarise the data through summarise function
```{r}
chicago <- mutate(chicago, year = as.POSIXlt(date)$year+1900)
head(chicago)
```

```{r}
years <- group_by(chicago, year)
summarise(years, pm25_m = mean(pm25, na.rm = T), 
          dp_avg = mean(dewpoint, na.rm = T) )
```
## Pipeline operator
The pipeline operator is %>% where you can pass the data frame through a pipeline
of operations. 
```{r}
chicago %>% mutate(month = as.POSIXlt(date)$mon+1) %>% group_by(
        month) %>% summarise(dp_mean = mean(dewpoint, na.rm=T), 
                             dp_median = median(dewpoint, na.rm = T),
                             pm25_ave = mean(pm25, na.rm=T)) 

```
## dplyr can work with 
* Work with data table package
* SQL interface for RDBMS






