---
title: "Reshaping Data"
author: "Sumit Kamra"
date: "8/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reshaping Data
#### Use library reshape2

```{r}
library(reshape2)
head(mtcars)
```
## Melting data frames
```{r}
mtcars$carname <- rownames(mtcars)
carMelt <- melt(mtcars, id=c("carname", "gear", "cyl"), 
                measure.vars = c("mpg", "hp") )
head(carMelt)
tail(carMelt)
```
Create bunch of id values and melt all the values - converting into a tall and 
skinny data set. 

## Reacsting the data 
Let's recast the data by gears by using the dcast function. 
```{r}
gearData <- dcast(carMelt, gear ~ variable)
gearData        
```
This tells us how many observations we have for mpg and hp for different gears
```{r}
gearData <- dcast(carMelt, gear ~ variable, mean)
gearData
```
## Averaging values
```{r}
head(InsectSprays)
```
*The function below applies to count, along sprays, the function sum* 
```{r}
tapply(InsectSprays$count, InsectSprays$spray, sum)
```
## Split, apply, combine
```{r}
spIns <- split(InsectSprays$count, InsectSprays$spray)
spIns
```

```{r}
sprCount <- lapply(spIns, sum)
sprCount
```
```{r}
unlist(sprCount)
```

## plyr function 
```{r}
library(plyr)
ddply(InsectSprays, .(spray), summarise, sum = sum(count), 
      average = round(mean(count),2))
```


## References
[Interesting tutorial for data reshaping](https://www.slideshare.net/jeffreybreen/reshaping-data-in-r)  

[R Blogger: A quick primer on split apply combine problems](https://www.r-bloggers.com/2011/12/a-quick-primer-on-split-apply-combine-problems/)

...



