---
title: "Subsetting and Sorting"
author: "Sumit Kamra"
date: "8/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subsetting
Let's start with loading airquality database. 

```{r loaddata}
data("airquality")
set.seed(10)
dat <- airquality[sample(1:153, 50),] 
head(dat)
```

###### Subsetting both rows and columns.
```{r subsetexamples}
dat[c(1,2,3), "Ozone"]

```
##### Subsetting using logical statement
```{r}
dat[(dat$Month==6 | dat$Month== 7),]
```

##### Subsetting on rows with NA (missing values) -- use which command
which command returns the indices 
```{r}
dat[which((dat$Month==6 | dat$Month== 7) & dat$Ozone <50),]
```

## Sorting values
##### Sort using sort function which sorts in a ascending order

```{r}
sort(dat$Ozone)
```
```{r}
sort(dat$Ozone, decreasing = T)
```
```{r}
sort(dat$Ozone, na.last = TRUE)
```
##### USe order function to sort a data frame.

```{r}
dat[order(dat$Month, dat$Temp),]
```

##### Use plyr package to sort the data frame
```{r}
library(plyr)
head(arrange(dat, Ozone), 5)

```
```{r}
head(arrange(dat, -Ozone))
```
##### Adding new columns to the data frame
```{r}
dat$TempCel <- round(((dat$Temp - 32) * 5 / 9), 1)
head(dat)
```
You can also do it through cbind and rbind


