---
title: "Summarizing Data"
author: "Sumit Kamra"
date: "8/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summarizing Data

```{r}
resData <- read.csv("Restaurants.csv")
head(resData, 3)
tail(resData, 3)
```
```{r}
summary(resData)
```
#### quantile function
```{r}
quantile(resData$Y, na.rm = T)
```

```{r}
quantile(resData$Y, probs = c(0, 0.5, 0.9), na.rm = T)
```
#### table function 
```{r}
table(resData$zipcode, useNA = "ifany")
```
#### Two dimensional table
```{r}
table(resData$zipcode, resData$cncldst, useNA = "ifany")
```
#### Count the numbers of NAs for a column

```{r}
sum(is.na(resData$zipcode))
```


```{r}
sum(is.na(resData$url))
```
#### Using any function to check if any of the values is na

```{r}
any(is.na(resData$prcl_pin))
```

#### all() function checks if every value satifies the condition

```{r}
all(resData$zipcode > 0)
```

```{r}
colSums(is.na(resData))
```

#### Value with specific characteritics 

```{r}
table(resData$zipcode %in% "21212")
```


```{r}
table(resData$zipcode %in% c("21212", "21213"))
```

#### use the %in% for subsetting
```{r}
head(resData[resData$zipcode %in% c("21212", "21213"), ], 3)

```

## Crosstabs 
```{r}
data("UCBAdmissions")
mydat <- as.data.frame(UCBAdmissions)
summary(mydat)
```
## Cross Tabs
#### Using xtabs function

```{r}
xtabs(Freq ~ Gender + Admit, data = mydat)
```
```{r}
xtabs(Freq ~ Dept + Admit, data = mydat)
```
#### Three dimensional xtabs and use of flat tabs
```{r}
xtabs(Freq ~ Gender + Admit + Dept, data = mydat)
```
#### This can also be written as . => means ALL 
```{r}
xt <- xtabs(Freq ~ ., data = mydat)
ftable(xt)
```
```{r}
fakeData <- rnorm(1e5)
object.size(fakeData)
print(object.size(fakeData), units = "MB")
```

